**NAME OF THIS STUDY**

### 2018-11-06 submission/manuscript.Rmd

# Read in OTU table and remove label and numOtus columns
shared <- read.delim('../data/baxter.0.03.subsample.shared', header=T, sep='\t') %>%
   select(-label, -numOtus)
```

```{r BinaryPreProcess, eval=TRUE, echo=FALSE, cache=FALSE}
# Merge metadata and OTU table and remove all the samples that are diagnosed with adenomas. Keep only cancer and normal.
# Then remove the sample ID column
data <- inner_join(meta, shared, by=c("sample"="Group")) %>%
  filter(dx != 'adenoma') %>%
  select(-sample)

# We want the diagnosis column to a factor
data$dx <- factor(data$dx, labels=c("normal", "cancer"))
```

```{r BinaryClassification, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
cl <- makePSOCKcluster(4)
registerDoParallel(cl)
# Create
all.test.response <- all.test.predictor <- test_aucs <- c()
for (i in 1:50) {
  inTraining <- createDataPartition(data$dx, p = .80, list = FALSE)
  training <- data[ inTraining,]
  testing  <- data[-inTraining,]
  x_train <- training %>% select(-dx)
  y_train <- training$dx
  y_train <- as.factor(y_train)
  # L2 regularization is used for Logistic regression
  # cost is the inverse of penalty and we provide a grid
  grid <-  expand.grid(cost = c(0.0000001, 0.000001, 0.00001, 0.0001),
                       loss = "L2_dual",
                       epsilon = 0.1)
  cv <- trainControl(method="repeatedcv",
                     repeats = 50,
                     number=5,
                     returnResamp="final",
                     classProbs=TRUE,
                     summaryFunction=twoClassSummary,
                     indexFinal=NULL,
                     preProc = "scale",
                     savePredictions = TRUE)

  L2Logit <- train(x_train, y_train,
                               method = "regLogistic",
                               trControl = cv,
                               metric = "ROC",
                               tuneGrid = grid,
                               preProcess = c("center", "scale"), 
                               family = "binomial")
  # Mean AUC value of the best lambda parameter training over repeats
  cv_auc <- getTrainPerf(L2Logit)$TrainROC
  # Predict on the test set and get predicted probabilities
  probs <- predict(L2Logit, testing, type="prob")
  # Test AUC calculation
  test_roc <- roc(ifelse(testing$dx == "cancer", 1, 0), probs[[2]])
  test_auc <- test_roc$auc
  # Save all the test AUCs over iterations in test_aucs
  test_aucs <- c(test_aucs, test_auc)
  # Save the test set labels in all.test.response. Labels converted to 0 for normal and 1 for cancer
  all.test.response <- c(all.test.response, ifelse(testing$dx == "cancer", 1, 0))
  # Save the test set predicted probabilities of highest class in all.test.predictor
  all.test.predictor <- c(all.test.predictor, probs[[2]])
}
stopCluster(cl)
# Get the ROC curve and AUC values for all the iterations 
test_roc <- roc(all.test.response, all.test.predictor, auc=TRUE, ci=TRUE)
```

```{r LogitROCplot, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

pdf("../results/figures/LogReg_inR.pdf")
par(mar=c(4,4,1,1))
# Plot random line on ROC curve
plot(c(1,0),c(0,1),
     type='l',
     lty=3,
     xlim=c(1.01,0), ylim=c(-0.01,1.01),
     xaxs='i', yaxs='i',
     ylab='', xlab='')
# Plot Test ROC in red line
plot(test_roc,
     col='black',
     lwd=2,
     add=T,
     lty=1)
#Calculate Confidence Interval sensitivities at given specificities
sens.ci <- ci.se(test_roc)
# Compute the CI of the AUC
auc.ci <- ci.auc(test_roc)
# Plot CV ROC in blue line
#plot(cv_roc,
#     col='blue',
#     lwd=2,
#     add=T,
#     lty=1)
# Label the axes
mtext(side=2,
      text="Sensitivity",
      line=2.5,
      cex=1.5)
mtext(side=1,
      text="Specificity",
      line=2.5,
      cex=1.5)
# Add legends for both lines
legend(x=0.7,y=0.2,
       legend=(sprintf('Test AUC: %.3g, CI: %.3g', test_roc$auc, auc.ci)),
       bty='n',
       xjust=0,
       lty=c(1,1),
       col='black',
       text.col='black')

plot(sens.ci, type="shape", col="gray88")

# Save the figure
dev.off()
```

>>>>

# Read in OTU table and remove label and numOtus columns
shared <- read.delim('../data/baxter.0.03.subsample.shared', header=T, sep='\t') %>%
   select(-label, -numOtus)

# Read in AUCs table generated from l2 logistic regression model for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMax scaler
logit <- read.delim('../data/process/L2_Logistic_Regression.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-Logistic Regression")

# Read in AUCs table generated from l1 SVM linear kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
l1svm <- read.delim('../data/process/L1_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L1-SVM Linear")

# Read in AUCs table generated from l2 SVM linear kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
l2svm <- read.delim('../data/process/L2_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-SVM Linear")

# Read in AUCs table generated from  SVM RBF kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
svmRBF <- read.delim('../data/process/SVM_RBF.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="SVM RBF")

# Read in AUCs table generated from xgboost for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
xgboost <- read.delim('../data/process/XGBoost.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="XGBoost")

# Read in AUCs table generated from random forest for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
rf <- read.delim('../data/process/Random_Forest.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Random Forest ")

# Read in AUCs table generated from decision tree for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
dt <- read.delim('../data/process/Decision_Tree.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Decision Tree")

# Put  all the AUCs for all the models together and group by model
all <- bind_rows(logit, l1svm, l2svm, svmRBF, xgboost, rf, dt) %>% 
  group_by(model)
```

```{r model_parameters, echo=FALSE, results='asis'}
model_param <- read.table('../data/process/model_parameters.txt', header=T, sep='\t')
kable(model_param, caption = "Table 1 Parameters and software implementation of the classification algorithms")
```
\vspace{35mm}
